### 2023-03-16 manuscript/main.tex
need tree then done

% !TeX document-id = {d42653b9-9ea4-48c2-860b-3d6e6e9d8ab0}
% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
\documentclass[12pt]{article} % For LaTeX2e
%\usepackage{neurips_2021}
\usepackage[colorlinks, citecolor={blue}]{hyperref}
\usepackage{url}
\usepackage{amsfonts,amscd,amssymb}
\usepackage{amsthm,amsmath,natbib}
\usepackage{algorithm,algorithmicx,algpseudocode}
\usepackage{bm}
\usepackage{bbm} %bb font numbers
\usepackage[table]{xcolor}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{natbib}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
%\usepackage[nolists]{endfloat}
\usepackage{listings}
\usepackage[textsize=tiny]{todonotes}
\usepackage{tikz}
\usetikzlibrary{shapes.misc}
\usepackage{etoolbox}
\usepackage{appendix}
\usepackage[format=plain,
labelfont={it},
textfont=it]{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{xr}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{mathbbol}
\usepackage{thmtools}
\usepackage{amsfonts}

\usetikzlibrary{matrix}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{arrows,shapes}
\usetikzlibrary{decorations}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{fit}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{shapes.geometric}

\newtoggle{quickdraw}
%\toggletrue{quickdraw} % Uncomment this to render more quickly (non-random)

\definecolor{lightgrey}{rgb}{0.9,0.9,0.9}
\definecolor{darkgreen}{rgb}{0,0.3,0}
%\definecolor{darkred}{rgb}{0.3,0,0}

\definecolorset{rgb}{}{}{darkred,0.8,0,0;darkgreen,0,0.5,0;darkblue,0,0,0.5}

%\doublespacing

\newtheorem{thm}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{mydef}{Definition}
\newtheorem*{assumption}{Assumption}
\newtheorem{clm}{Claim}

\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand*{\fplus}{\genfrac{}{}{0pt}{}{}{+}}
\newcommand*{\fdots}{\genfrac{}{}{0pt}{}{}{\cdots}}
\newcommand{\mb}{\mathbf}
\newcommand{\mc}{\mathcal}
\newcommand{\dx}{\mbox{d}}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\numTaxa}{N}
\newcommand{\numTraits}{D}
\newcommand{\numDatasets}{M}
%\newcommand{\numLatent}{D}
\newcommand{\taxonIndex}{i}
\newcommand{\traitIndex}{j}
\newcommand{\traitData}{\vec{Y}}
\newcommand{\traitDatum}{y}
\newcommand{\datasetIndex}{m}
\newcommand{\exemplar}{\text{e}}

\newcommand{\sequences}{\vec{S}}
\newcommand{\latentData}{\vec{X}}
\newcommand{\latentdata}{\vec{x}}
\newcommand{\latentDatum}{x}
\newcommand{\phylogeneticParameters}{\boldsymbol{\phi}}
\newcommand{\phylogeny}{{\cal G}}
\newcommand{\tree}{\phylogeny}
%\newcommand{\otherParameters}{\boldsymbol{\
\newcommand{\transpose}{^{t}}

\newcommand{\distanceMatrix}{\mathbf{Y}}
\newcommand{\distance}{y}
\newcommand{\summant}{r}

\newcommand{\cdensity}[2]{\ensuremath{p(#1 \,|\,#2)}}
\newcommand{\density}[1]{\ensuremath{p(#1 )}}

\newcommand{\treeNode}{\nu}

\newcommand{\traitVariance}{\mathbf{\Sigma}}
\newcommand{\nodeIndex}{c}

%\newcommand{\parent}[1]{\mbox{\tiny pa}(#1)}
\newcommand{\parentBig}[1]{\mbox{pa}(#1)}

\newcommand{\sibling}[1]{\mbox{\tiny sib}(#1)}
\newcommand{\siblingBig}[1]{\mbox{sib}(#1)}

\newcommand{\rootMean}{\boldsymbol{\mu}_0}
\newcommand{\rootVarianceScalar}{\tau_0}
\newcommand{\unsequencedVarianceScalar}{\tau_{\exemplar}}
\newcommand{\treeVariance}{\vec{V}_{\tree}}
\newcommand{\hatTreeVariance}{\hat{\vec{V}}_{\tree}}
\newcommand{\mdsSD}{\sigma}
\newcommand{\mdsVariance}{\mdsSD^2}
\newcommand{\residual}{\hat{\traitDatum}}
\newcommand{\modelDistance}{\delta}
\newcommand{\cdf}{\phi}
\newcommand{\normalCDF}[1]{\Phi \left( #1 \right)}

\newcommand{\order}[1]{{\cal O}\hspace{-0.2em}\left( #1 \right)}

\newcommand{\rootNode}{\nu^{\datasetIndex}_{2 \numTaxa_{\datasetIndex} -1 }}
\newcommand{\pathLength}[1]{d(F, #1 )}
\newcommand{\pathLengthNew}[2]{
d_{F}
(
{#1}, {#2}
)
}
\newcommand{\J}{\vec{J}}
\newcommand{\pprime}{^{\prime}}
\newcommand{\otherIndex}{i \pprime}
\def\kronecker{\raisebox{1pt}{\ensuremath{\:\otimes\:}}}

\definecolor{trevorblue}{rgb}{0.330, 0.484, 0.828}
\definecolor{trevoryellow}{rgb}{0.829, 0.680, 0.306}

\allowdisplaybreaks

%\makeatletter
%\def\title@font{\Huge}
%\let\ltx@maketitle\@maketitle
%\def\@maketitle{\bgroup%
%	\let\ltx@title\@title%
%	\def\@title{\resizebox{\textwidth}{!}{%
%			\mbox{\title@font\ltx@title}%
%	}}%
%	\ltx@maketitle%
%	\egroup}
%\makeatother

\title{Approximating the matrix exponential derivative of the CTMC generator matrix, with application to global SARS-CoV-2}

%\date{}

\author[1]{Andrew J.~Holbrook}
\author[1]{Andrew F.~Magee}
\author[2]{Nathan E.~Glatt-Holtz}
\author[1,3,4]{Marc A.~Suchard}

 \affil[1]{Department of Biostatistics, University of California, Los Angeles}
 \affil[2]{Department of Mathematics, Tulane University}
 \affil[3]{Department of Biomathematics, University of California, Los Angeles}
 \affil[4]{Department of Human Genetics, University of California, Los Angeles}

 \renewcommand\Authands{ and }

\graphicspath{{figures/}}

\begin{document}

\maketitle

\begin{abstract}

We investigate the accuracy of a first-order approximation to the directional derivative of the matrix exponential evaluated on continuous-time Markov chain (CTMC) generator matrices.

\end{abstract}

\section{Introduction}\label{sec:intro}

%\newcommand{\x}{\mathbf{x}}
%\renewcommand{\v}{\mathbf{v}}

%\newcommand{\TTTheta}{\boldsymbol{\Theta}}
%\newcommand{\tr}{\mbox{tr}}
%\newcommand{\X}{\mathbf{X}}
%\renewcommand{\u}{\mathbf{u}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\QQ}{\mathbf{Q}}
\newcommand{\DD}{\mathbf{D}}
\newcommand{\MM}{\mathbf{M}}
\newcommand{\JJ}{\mathbf{J}}
\newcommand{\II}{\mathbf{I}}
\newcommand{\RR}{\mathbf{R}}
%\newcommand{\haar}{\mathcal{H}}
%\newcommand{\orthog}{\mathcal{O}}
%\newcommand{\SSigma}{\boldsymbol{\Sigma}}
\newcommand{\Zero}{\boldsymbol{0}}
\newcommand{\ttheta}{\boldsymbol{\theta}}
\newcommand{\Llambda}{\boldsymbol{\Lambda}}
\newcommand{\vecc}{\mbox{vec}}
\newcommand{\CC}{\mathbf{C}}
\newcommand{\BB}{\mathbf{B}}
\renewcommand{\AA}{\mathbf{A}}
\newcommand{\GG}{\mathbf{G}}
\newcommand{\mom}{\boldsymbol{\xi}}
\newcommand{\ppi}{\boldsymbol{\pi}}
\renewcommand{\P}{\mathbf{P}}
\newcommand{\p}{\mathbf{p}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\X}{\mathbf{X}}

 We are interested in efficiently approximating the directional derivative of the matrix exponential
\begin{align*}
	e^{t\QQ} = \sum_{k=0}^\infty \frac{t^k\QQ^k}{k!}
\end{align*}
for arbitrary $t>0$ using the formula \citep{najfeld1995derivatives}
\begin{align}\label{eq:deriv}
	\nabla_{\JJ} e^{t\QQ}  =  e^{t\QQ}  \sum_{k=0}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} \, .
\end{align}
Here, $\JJ$ gives the direction and $\{\JJ,\QQ^k\}$ is the matrix commutator series satisfying the recursion
\begin{align*}
	\{\JJ,\QQ^0\} &= \JJ \, , \\ 
	\{\JJ,\QQ^k\} &= [\{\JJ,\QQ^{k-1}\} , \QQ]  = \{\JJ,\QQ^{k-1}\} \QQ - \QQ \{\JJ,\QQ^{k-1}\} \, .
\end{align*}
Thus, the first-order approximation to derivative \eqref{eq:deriv} is 
\begin{align}\label{eq:firstOrder}
	\widetilde{\nabla}_{\JJ} e^{t\QQ}= t e^{t\QQ}\JJ \, .
\end{align}

\section{Global spread of SARS-CoV-2}

\subsection{Surrogate trajectory HMC}\label{sec:surrHmc}

Hamiltonian Monte Carlo (HMC) \citep{duane1987hybrid,neal2011mcmc} is an advanced MCMC procedure that uses numerical approximations to Hamiltonian trajectories to generate Metropolis-type \citep{metropolis1953equation} proposals far away from the current Markov chain state.  On the one hand, this approach to proposal generation helps reduce autocorrelation between chain states and is particularly helpful within higher-dimensional state spaces.  On the other hand, numerical integration of Hamilton's equations requires repeated evaluation of the Hamiltonian potential energy's gradient, and these repeated floating-point operations may become computationally burdensome.   Using HMC for Bayesian inference, the potential energy takes the form $u(\ttheta)=-\log \pi(\ttheta)$, where $\pi(\cdot)$ is the posterior density function $\pi(\ttheta):=p(\ttheta|\y)$ of parameter $\ttheta\in\mathbb{R}^D$ given observed data $\y$.  In this context, a single iteration of the the St\"{o}rmer-Verlet (velocity Verlet) or leapfrog integrator \citep{leimkuhler2004simulating}, beginning at time $\tau$ and with $\epsilon>0$ small, takes the form
\begin{align}\label{eq:leap}
	\mom\left(\tau + \frac{\epsilon}{2}\right) &= \mom(\tau) + \frac{\epsilon}{2} \nabla \log \pi(\ttheta(\tau))  \\  \nonumber
	\ttheta(\tau+\epsilon) &= \ttheta(\tau) + \epsilon \, \MM^{-1}\mom(\tau+\frac{\epsilon}{2}) \\ \nonumber
	\mom(\tau + \epsilon) &= \mom\left(\tau+\frac{\epsilon}{2}\right) + \frac{\epsilon}{2} \nabla \log \pi(\ttheta(\tau+\epsilon))  \, ,
\end{align}
where $\mom \sim \mbox{N}_D(\Zero,\MM)$ is the Hamiltonian system's Gaussian-distributed momentum variable.  
Different strategies aim to speed up the leapfrog integrator's many log-posterior gradient evaluations $\nabla \log \pi (\ttheta)$ in different ways, as these numerical routines often represent the algorithm's computational bottleneck.  In model-specific contexts, \citet{holbrook2021massive,holbrook2022bayesian,holbrook2022viral} develop parallelization strategies, and \citet{ji2020gradients} develop dynamic programming techniques, to accelerate $\nabla \log \pi (\ttheta)$.

A small body of work considers another approach by replacing $\nabla \log \pi (\ttheta)$ with a suitable approximation $\widehat{\nabla} \log \pi (\ttheta)$ and recognizing that the modified \eqref{eq:leap} continues to satisfy path reversibility and volume preservation, two essential ingredients for well-specified HMC. It follows immediately that the resulting `surrogate trajectory HMC' continues to sample the correct target distribution $\pi(\cdot)$.  Nonetheless, the acceptance rates and overall efficiency of such samplers may suffer when approximations are poor.   The majority of surrogate HMC methods first obtain a small sample of exact gradient evaluations and then use some model to interpolate: \citet{zhang2017precomputing} assume the approximate gradient follows a piecewise constant form across a grid; \citet{rasmussen2003gaussian,lan2016emulation} construct approximations using Gaussian processes; and \citet{zhang2017hamiltonian,li2019neural} do the same using neural networks.  But an even simpler approach to surrogate HMC may be appropriate when gradients have series representations.   \textcolor{red}{MAGEE} consider phylogenetic models that involve CTMC priors and show that the first-order approximation \eqref{eq:firstOrder} of the matrix exponential derivative \eqref{eq:deriv} performs remarkably well within surrogate HMC, achieving an over 30-fold efficiency gain compared to random-walk Metropolis for a 14-state model with over 180 model parameters.  Here, we demonstrate similar strong performance of the first-order approximation within surrogate HMC for a 44-state model with almost 1,900 model parameters.

\subsection{Phylogenetic CTMC model}

Start with a (possibly unknown) rooted and bifurcating phylogenetic tree $\phylogeny$ consisting of $N$ leaf nodes that correspond to observed biological specimens and $N-1$ internal nodes that correspond to unobserved ancestors.  The tree also contains $2N-2$ branches of length $t_v$ connecting each child node $v$ to its parent $u$.  
 Given $\phylogeny$, we model the evolution of $S$ characters along each branch of the tree according to a CTMC model with $S\times S$ generator matrix $\QQ$ and stationary distribution $\ppi=(\pi_1,\dots,\pi_S)=\lim_{t\rightarrow \infty}e^{t\QQ}\widetilde{\ppi}$, for $\widetilde{\ppi}$ for any arbitrary probability vector.   Examples of characters are the $S=4$ nucleotides within a set of aligned genome sequences  \citep{jukes1969evolution} and the set of $S=15$ geographic regions visited by an influenza subtype \citep{lemey2014unifying}.   We may scale $t_v$ to be raw time (e.g., years in Section \ref{sec:sars}) or the expected number of substitutions with respect to $\ppi$ depending on the given problem.   In the former case, one may augment the model with a rate scalar $\gamma$ that modulates the expected number of substitutions across all branches, and the finite-time transition probability matrix along branch $v$ becomes $\P_v:=e^{\gamma t_v \QQ}$.  In the following, we further posit $\QQ=\QQ(\ttheta)$ for $\ttheta$ a vector of parameters.

 Let data $\y$ be the $N$-vector with elements $y_n\in \{1,\dots,S\}$, the observed states of each observed biological specimen $n \in \{1,\dots,N\}$.  At any node $v$, the likelihood takes the form \citep{ji2020gradients}
 \begin{align}
p(\y|\ttheta) = \p_v^T \q_v \, ,
 \end{align}
where $\p_v$ and $\q_v$ are the post-order and pre-order partial likelihood vectors, respectively.  The former describes the probability of the observed states for all observed specimens (i.e., leaf nodes) that descend from node $v$, conditioned on the state at node $v$. The latter describes analogous probabilities for all observed specimens not descending from node $v$.  If we suppose that node $u$ gives rise to two child nodes, $v$ and $w$, then 
\begin{align}\label{eq:recursions}
	\p_u = \P_v \p_v \circ \P_w \p_w\, ,  \quad \q_v= \P_v^T \left( \q_u \circ  \P_w \p_w\right) \, .
\end{align}
Using the chain rule and the fact that $\p_v$ does not depend on $\P_v$, one may write the likelihood's derivative with respect to a single parameter $\theta_d$ thus:
\begin{align}\label{eq:phyloGrad}
	\frac{\partial}{\partial \theta_d}  p(\y|\ttheta) 
	&=   \sum_{v=1}^{2N-2} \mbox{tr} \left( \frac{\partial (\p_v^T\q_v)}{\partial \P_v} \frac{\partial \P^T_v}{\partial \theta_d} \right)
	= \sum_{v=1}^{2N-2} \mbox{tr} \left(  (\q_u\circ \P_w\p_w) \p_v^T\left( \frac{\partial e^{\gamma t_v \QQ}}{\partial \theta_d}\right)^T\right)  \\ \nonumber
	&=\sum_{v=1}^{2N-2} \p_v^T \left(\sum_{s=1}^S \sum_{s'=1}^S \frac{\partial e^{\gamma t_v \QQ}}{\partial q_{ss'}} \frac{\partial q_{ss'}}{\partial \theta_d} \right)^T (\q_u\circ \P_w\p_w)  \, .
\end{align}
Whereas the recursions of \eqref{eq:recursions} facilitate fast likelihood computation, inferring $\ttheta$ using gradient-based approaches such as HMC (Section \ref{sec:surrHmc}) requires a large number of repeated evaluations of the matrix exponential derivative.  These computations become particularly onerous when one opts for a gold-standard mixed-effect model \textcolor{red}{(MAGEE)} and specifies
\begin{align}\label{eq:mixed}
\log q_{ss'} = b_{ss'} + \epsilon_{ss'} \,, \quad s\neq s'\, .
\end{align} 
Here, the fixed-effects $b_{ss'}$ are elements of some non-random matrix $\B$, and the random-effects $\epsilon_{ss'}$ are mutually independent \emph{a priori} and inferred as model parameters.  The dimension $D$ of $\ttheta$ in this model is $\mathcal{O}(S^2)$, so the usual $\mathcal{O}(DNS^5)$ cost of the matrix expoential derivative becomes a massive $\mathcal{O}(NS^7)$.  In this context, \textcolor{red}{MAGEE} show that the first-approximation \eqref{eq:firstOrder} achieves a considerable speedup over the exact derivative, while facilitating high-quality proposals in the context of surrogate HMC.  In the following section, we use this method to analyze the global spread of SARS-CoV-2 and show that the first-order approximation maintains its performance for an even higher-dimensional problem than previously considered.

\subsection{Bayesian analysis of SARS-CoV-2 contagion}\label{sec:sars}

\begin{figure}[!t]
	\centering
	\includegraphics[width=\linewidth]{fixedEffects.pdf}
	\caption{Fixed effects.}\label{fig:FE}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=\linewidth]{postViz.pdf}
	\caption{Posteriors.}\label{fig:post}
\end{figure}

We use the phylogenetic CTMC framework to model the early spread of SARS-CoV-2---the virus responsible for the ongoing COVID-19 pandemic---based on 285 observed viral samples collected from $S=44$ regions worldwide. These regions comprise 15 provinces within China and 29 countries without.  Understanding the manner in which viruses travel between human populations is an object of ongoing study, and phylogeographic analyses point to the central role of travel networks including those measured by airline passenger counts \citep{holbrook2021massive} or google mobility data \citep{worobey2020emergence}.  Here, we include three such predictors of travel in our CTMC model by expanding the fixed-effects in regression model \eqref{eq:mixed} to take the form
\begin{align*}
	\B = \theta_1 \X_1 + \theta_2  \X_2 + \theta_3 \X_3 \, ,
\end{align*} 
for each $\X$ a fixed $44\times 44$ matrix predictor and each $\theta$ a real-valued regression coefficient. Figure \ref{fig:FE} presents the three predictors of interest: $\X_1$ contains air travel proximities between locations as measured by annualized air passenger counts between airports contained within a region \citep{holbrook2021massive};  $\X_2$ contains intracontinental proximities arising from physical distances when two regions inhabit the same continent and fixed at the minimum otherwise; finally, $\X_3$ describes the Hubei asymmetry, i.e., the fact that people could enter, but not leave, the Chinese province of Hubei during the early pandemic.

In the context of a Bayesian analysis, we specify independent normal priors on $\theta_1$, $\theta_2$ and $\theta_3$ with means of 0 and variances of 2.  We also assume that the 1,892 random-effects $\epsilon_{ss'}$ follow sparsity inducing Bayesian bridge priors with global scale parameter $\tau$ and exponent $\alpha=0.25$.  Here, we follow \citet{nishimura2022shrinkage} and specify a Gamma prior on $\tau^{-\alpha}$ with a shape parameter of 1 and a rate parameter of 2. Finally, we place a flat prior on the rate scalar $\gamma$.  Inferring the posterior distribution of all 1,897 model parameters requires an advanced MCMC strategy. Namely, we adopt an HMC-within-Gibbs approach, updating the scalars $\gamma$ and $\tau$ independently but updating all 1,895 regression parameters (both fixed- and random-effects) using surrogate HMC accomplished with the first-order approximation
\begin{align*}
  \gamma t_v  e^{\gamma t_v \QQ} \JJ_{ss'}    \approx \frac{\partial e^{\gamma t_v \QQ}}{\partial q_{ss'}} 
\end{align*}
within \eqref{eq:phyloGrad}.

We generate 8 million MCMC samples in this manner, saving 1 in 10,000 Markov chain states, in order to guarantee a minimum effective sample size (ESS) greater than 100.  By far, the worst-mixing parameter  is the global scale $\tau$, which obtains an ESS of 185.  The three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ obtain ESS's of 721, 721 and 644, respectively.  The median and minimum ESS for the 1,892 random-effects $\epsilon_{ss'}$ are 721 and 190, respectively.  We note that, after thinning and removing burn-in, the sample only consists of 721 states, so ESS of 721 implies negligible autocorrelation between samples.  The left plot of Figure \ref{fig:post} presents posterior densities for the fixed-effect coefficients from regression \eqref{eq:mixed}.  The air traffic proximity, intracontinental proximity and Hubei asymmetry coefficients have posterior means and 95\% credible intervals of 0.76 (0.50, 1.03), 0.04 (-0.04, 0.12) and 0.27 (-0.39, 0.78), respectively. While these results suggest a positive association between each predictor matrix and the infinitesimal generator matrix $\QQ$, the only predictor with a statistically significant association is air traffic proximity.  This result agrees with a previous phylogeographic analysis of the global spread of influenza \citep{holbrook2021massive}.  The right plot of Figure \ref{fig:post} presents the posterior mean for generator $\QQ$.  As one may expect, the matrix looks similar to that of the air traffic predictor matrix in Figure \ref{fig:FE}, but one may also see the influence of the Hubei asymmetry in, e.g., the squares corresponding to travel between Guangdong and Hubei provinces.

\begin{itemize}
	\item tree
\end{itemize}

In addition to these scientific questions of interest, we are interested in the performance of the first-order approximation as a surrogate gradient for HMC in such a high-dimensional setting. To isolate this performance, we fix the Bayesian bridge global-scale $\tau$ at 2.5$\times10^{-5}$.  We generate a Markov chain with 80,000 states, saving every tenth state and removing the first 1000 states as burn-in.  Despite the relatively small number of iterations, we observe large ESS that suggest satisfactory accuracy of the first-order approximation within the context of high-dimensional surrogate HMC.  The ESS for the three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ are 1,053, 1,343 and 498, respectively.  The median and minimum ESS for the 1,892 random-effects $\epsilon_{ss'}$ are 1,514 and 1,161, respectively.

\section{Discussion}

\section*{Acknowledgments}

AJH is supported by a gift from the Karen Toffler Charitable Trust and by grants NIH K25 AI153816, NSF DMS 2152774 and NSF DMS 2236854.

%Let $\QQ=[q_{ij}]$ be the rate matrix for an $S$-dimensional continuous time Markov chain, where $q_{ij}>0$ for $i\neq j$, and $q_{ii} = -\sum_{j\neq i} q_{ij}$.  In our application of interest, $\QQ$ is almost surely diagonalizable, i.e., $\QQ=\MM \DD \MM^{-1}$ for 
%\begin{align*}
%	\DD = \begin{pmatrix}
%		d_1& 0 &\dots & & \\
%		0  & d_2 &0  & \dots& \\
%		\vdots &0 & \ddots & \\
%		& \vdots& & d_{S-1} &0 \\
%		&&&0&0
%	\end{pmatrix} \, ,
%\end{align*}
%where the structure of $\QQ$ implies eigenvalues
%\begin{enumerate}
%	\item $d_s\in \mathbb{C}$ with $\mbox{Re}(d_s)<0$ for $s\in\{1,\dots,S-1\}$, and
%	\item $d_S =0$ corresponding to the eigenvector $\boldsymbol{1}=(1,\dots,1)^T$. 
%\end{enumerate} 
%We may further define a generalized inverse $\QQ^+=\MM \DD^{+} \MM^{-1}$ of $\QQ$ using $\DD^+$, the Moore-Penrose pseudoinverse obtained by inverting the non-zero elements of $\DD$.   Finally, we note that diagonalizability implies $\DD$ and $\DD^+$ commute, i.e., $\DD^+\DD=\DD\DD^+$.
%
%
% We are interested in efficiently approximating the directional derivative of the matrix exponential
%\begin{align*}
%	e^{t\QQ} = \sum_{k=0}^\infty \frac{t^k\QQ^k}{k!}
%\end{align*}
%for arbitrary $t>0$ using the formula \citep{najfeld1995derivatives}
%\begin{align}\label{eq:deriv}
%	\nabla_{\JJ} e^{t\QQ}  =  e^{t\QQ}  \sum_{k=0}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} \, .
%\end{align}
%Here, $\JJ$ gives the direction and $\{\JJ,\QQ^k\}$ is the matrix commutator series satisfying the recursion
%\begin{align*}
%	\{\JJ,\QQ^0\} &= \JJ \, , \\ 
%	\{\JJ,\QQ^k\} &= [\{\JJ,\QQ^{k-1}\} , \QQ]  = \{\JJ,\QQ^{k-1}\} \QQ - \QQ \{\JJ,\QQ^{k-1}\} \, .
%\end{align*}
%Thus, the first-order approximation to derivative \eqref{eq:deriv} is 
%\begin{align}\label{eq:firstOrder}
%	\widetilde{\nabla}_{\JJ} e^{t\QQ}= t e^{t\QQ}\JJ \, .
%\end{align}
%We want to bound the truncation error
%\begin{align*}
%\mbox{err}(t,\QQ,\JJ)	:=\lVert \nabla_{\JJ} e^{t\QQ} - \widetilde{\nabla}_{\JJ} e^{t\QQ} \rVert \, ,
%\end{align*}
% where $\lVert \cdot \rVert$ denotes an appropriate matrix norm.  In this context, we find that naive bounds (Appendix \ref{sec:loose}) that do not exploit the structure of $\QQ$  are rather loose (Figure \ref{fig:frob}).  
% In our efforts to fully leverage the structure of $\QQ$, we will make use of the general formula for the matrix commutator series  \citep{volkin1968iterated}
%\begin{align}\label{eq:commSeries}
%	\{ \JJ, \QQ^k\} = \sum_{r=0}^k (-1)^{k+r}  {k \choose r} \QQ^{k-r} \JJ \QQ^r  \, .
%\end{align}
%
%\subsection{An empirically tight asymptotic estimate}
%
%We begin by noting
%\begin{align}\label{eq:ub1}
%	\mbox{err}(t,\QQ,\JJ) &=  \lVert  e^{t\QQ}  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} \rVert \\ \nonumber
%	&\leq  \lVert  e^{t\QQ}\JJ  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \QQ^k \rVert +   \lVert  e^{t\QQ}  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} -\JJ\QQ^k\rVert \\ \nonumber
%		&=  \lVert  e^{t\QQ}\JJ \QQ^+  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \QQ^{k+1} \rVert +   \lVert  e^{t\QQ}  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} -\JJ\QQ^k\rVert \\ \nonumber
%		&=  \lVert  e^{t\QQ}\JJ \QQ^+ (e^{t\QQ} - t\QQ -\II) \rVert +   \lVert  e^{t\QQ}  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} -\JJ\QQ^k\rVert \\ \nonumber
%			&\leq \lVert  e^{t\QQ}\JJ \QQ^+ ( t\QQ +\II) \rVert  + \lVert  e^{t\QQ}\JJ \QQ^+ e^{t\QQ} \rVert  +   \lVert  e^{t\QQ}  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} -\JJ\QQ^k\rVert \\ \nonumber
%			&=: \mathfrak{a} + \mathfrak{b} + \mathfrak{c} \, .
%\end{align}
%Empirical experiments using the Frobenius and infinity norms (Figures \ref{fig:frob} and \ref{fig:inf}, respectively) show that the first term
%\begin{align}
%	 \mathfrak{a} = \left\Vert e^{t\QQ} \JJ \QQ^+(t\QQ+\II)  \right\Vert \, ,
%\end{align}
%although not a strict upper bound, is asymptotically tight for the true error as $t\rightarrow \infty$. We would, therefore, like to show that the terms $\mathfrak{b}$ and $\mathfrak{c}$ go to zero as $t$ diverges.
%
%
%
%
%
%
%
%
%\section{Using the structure of $\QQ$}
%
%We use the fact that
%\begin{align*}
%	e^{t\QQ}\QQ = \MM \exp(t\DD) \DD \MM^{-1} =  \MM \begin{pmatrix}
%	e^{td_1}	d_1& 0 &\dots & & \\
%		0  & e^{td_2}d_2 &0  & \dots& \\
%		\vdots &0 & \ddots & \\
%		& \vdots& & e^{td_{S-1}}d_{S-1} &0 \\
%		&&&0&0
%	\end{pmatrix} \MM^{-1} \, .
%\end{align*}
%Thus, assuming use of a submultiplicative norm leads to the result
%\begin{align*}
%	\lVert e^{t\QQ}\QQ \rVert &\leq \left\Vert  \begin{pmatrix}
%		e^{td_1}	d_1& 0 &\dots & & \\
%		0  & e^{td_2}d_2 &0  & \dots& \\
%		\vdots &0 & \ddots & \\
%		& \vdots& & e^{td_{S-1}}d_{S-1} &0 \\
%		&&&0&0
%	\end{pmatrix}  \right\Vert  \lVert \MM\rVert   \lVert \MM^{-1}\rVert  \, .
%\end{align*}
%For, e.g., the Frobenius norm, we have
%\begin{align*}
%\lVert e^{t\QQ}\QQ \rVert_F	\leq \sqrt{S-1} \max_s \{|e^{td_{s}} d_{s}| \}  \lVert \MM\rVert_F   \lVert \MM^{-1}\rVert_F  \, ,
%\end{align*}
%where we  use the fact that $\lVert \cdot \rVert_F \leq \sqrt{r} \lVert \cdot \rVert_2$ for any rank $r$ matrix and $\lVert \cdot \rVert_2$ the spectral norm.  We therefore write
%\begin{align}
%	\lVert e^{t\QQ}\QQ \rVert_F = \order{\max_s \left\{|e^{td_{s}} d_{s}| \right\}} \,,  \quad 	\lVert e^{t\QQ}\QQ^+ \rVert_F = \order{\max_s \left\{ \left| e^{td_{s}} \frac{1}{d_{s}}\right| \right\}} \, .
%\end{align}
%Because $\mbox{Re}(d_s)<0$ for $s\in \{1,\dots,S-1\}$, we have
%\begin{align}
%	\lim_{t\rightarrow \infty}  	\lVert e^{t\QQ}\QQ \rVert_F = 	\lim_{t\rightarrow \infty}  	\lVert e^{t\QQ}\QQ^+ \rVert_F = 0\, ,
%\end{align}
%and, by equivalence of norms, we have
%\begin{align}
%	\lim_{t\rightarrow \infty}  	\lVert e^{t\QQ}\QQ \rVert = 	\lim_{t\rightarrow \infty}  	\lVert e^{t\QQ}\QQ^+ \rVert = 0\, ,
%\end{align}
%in general for any matrix norm $\lVert \cdot \rVert$. Moreover, the matrix $e^{t\QQ}$ is a stochastic matrix with rows summing to unity. Therefore $\lVert e^{t\QQ}\rVert_\infty = 1$, and, by norm equivalence, $\lVert e^{t\QQ}\rVert = \mathcal{O}(1)$.
% These results lead to the following fact for term $\mathfrak{b}$ in the upper bound of \eqref{eq:ub1}:
%\begin{align*}
%	\lim_{t\rightarrow \infty} \lVert  e^{t\QQ}\JJ \QQ^+ e^{t\QQ} \rVert \leq  \lim_{t\rightarrow \infty} \lVert  e^{t\QQ} \rVert \lVert \JJ  \rVert \lVert \QQ^+ e^{t\QQ} \rVert = 0 \, .
%\end{align*}
%Thus, we obtain the intermediary result for $t$ large:
%\begin{align*}
%\lim_{t\rightarrow \infty}\mbox{err}(t,\QQ,\JJ) \leq \lim_{t\rightarrow \infty}  \mathfrak{a}+ \mathfrak{c} \, .
%\end{align*}
%It remains to show that $\mathfrak{c}$ goes to zero as $t$ diverges, i.e., 
%\begin{align}
%	\lim_{t\rightarrow \infty} \lVert  e^{t\QQ}  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} -\JJ\QQ^k\rVert = 0\, .
%\end{align}
%
%
%
%\section{Random thoughts}
%
%\subsection{Another form}
%
%\newcommand{\PPsi}{\boldsymbol{\Psi}}
%
%\citet{najfeld1995derivatives} use the integral representation to prove the interesting result
%\begin{align*}
%	\nabla_{\JJ} e^{t\QQ} = \MM \left(  \left(\MM^{-1} \JJ \MM\right) \circ \PPsi(t)  \right) \MM^{-1} \, ,
%\end{align*}
%where $\PPsi(t)$ has elements
%\begin{align*}
%		[\PPsi(t)]_{ij} = \bigg\{ \begin{array}{cc}
%		t e^{td_i}\, ,	&  d_i = d_j \\
%		\frac{e^{td_i}-e^{td_j}}{d_i-d_j} \, ,& d_i \neq d_j 
%	\end{array} \, .
%\end{align*}
%For our application, the eigenvalues equal one another with measure zero.  Taking $t$ large, we have
%\begin{align*}
%	[\PPsi(t)]_{ij} = \bigg\{ \begin{array}{cc}
%		t e^{td_i}\, ,	&  i = j \\
%		0 \, ,& i \neq j 
%	\end{array} \, .
%\end{align*}
%It follows that, again for $t$ large,
%\begin{align*}
%	\mbox{err}(t,\QQ,\JJ) &\approx \lVert   \MM \left(  \left(\MM^{-1} \JJ \MM\right) \circ t e^{t\DD}  \right) \MM^{-1}       -  t e^{t\QQ} \JJ            \rVert \\
%	&= t\lVert   \MM \rVert \lVert \left(  \left(\MM^{-1} \JJ \MM\right) \circ  e^{t\DD}  \right) \MM^{-1}       -   e^{t\DD}\MM^{-1} \JJ            \rVert \, .
%\end{align*}
%
%\appendix
%
%\section{Loose bounds}\label{sec:loose}
%
%The following loose bounds do not leverage the structure of the CTMC generator matrix $\QQ$.
%
%\subsection{Upper bound 1}
%
%
%
%For a submultiplicative matrix norm, we have
%\begin{align*}
%	\Vert \{\MM_1, \MM_2^k \}  \Vert \leq 2^k \Vert \MM_1\Vert \Vert\MM_2 \Vert ^k \, .
%\end{align*}
%We use this to obtain our first upper bound on $err_{ij}(\QQ)$:
%\begin{align*}
%	err_{ij}(\QQ) &= \left\Vert e^{t\QQ}  \sum_{k=1}^\infty \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} \right\Vert  \\
%	&\leq \Vert e^{t\QQ} \Vert \sum_{k=1}^\infty \left\Vert \frac{t^{k+1}}{(k+1)!} \{\JJ,\QQ^k\} \right\Vert  \\
%	&=  \Vert e^{t\QQ} \Vert \sum_{k=1}^\infty \frac{t}{k+1} \frac{(2t \Vert \QQ \Vert)^k}{k!}\\
%	&= \Vert e^{t\QQ} \Vert   \frac{\exp(2t\Vert \QQ \Vert)-2t\Vert\QQ\Vert-1}{2\Vert \QQ \Vert} 
%\end{align*}
%
%
%\subsection{Upper bound 2}
%
%We start with another representation of the derivative \citep{al2009computing}:
%\begin{align*}
%	\nabla_{\JJ} e^{t\QQ} = \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^k \QQ^{p-1} \JJ \QQ^{k-p} 
%\end{align*}
%and the fact that
%the fact that, for the Frobenius norm $\Vert \cdot \Vert_F$, we have \citep{wu2010short}
%\begin{align*}
%	\Vert [\MM_1,\MM_2] \Vert_F \leq \sqrt{2} \Vert \MM_1 \Vert_F \Vert \MM_2\Vert_F \, .
%\end{align*}
%Using these results, the truncation error may be rewritten
%\begin{align*}
%	err_{ij}(\QQ)  &=\left\Vert - te^{t\QQ} \JJ  +  \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^k \QQ^{p-1} \JJ \QQ^{k-p}\right\Vert_F \\
%	&= \left\Vert \sum_{k=1}^\infty \frac{t^k}{k!} \left( \left( \sum_{p=1}^k \QQ^{p-1} \JJ \QQ^{k-p} \right) - k \QQ^{k-1} \JJ \right)  \right\Vert_F\\
%	&= \left\Vert \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^k \QQ^{p-1} \JJ \QQ^{k-p}  -  \QQ^{k-1} \JJ \right\Vert_F \\
%	&=\left\Vert  \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^{k-1} \QQ^{p-1} \JJ \QQ^{k-p}  -  \QQ^{k-1} \JJ\right\Vert_F \\
%	&=\left\Vert  \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^{k-1} \QQ^{p-1}\left( \JJ \QQ^{k-p}  -  \QQ^{k-p} \JJ \right)\right\Vert_F \\
%	&= \left\Vert  \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^{k-1} \QQ^{p-1}[\JJ, \QQ^{k-p}] \right\Vert_F  \\
%	&\leq  \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^{k-1} \Vert \QQ^{p-1} \Vert_F \Vert[\JJ, \QQ^{k-p}]\Vert_F 
%	\leq \sqrt{2}  \sum_{k=1}^\infty \frac{t^k}{k!} \sum_{p=1}^{k-1} \Vert \QQ^{k-1} \Vert_F  \\
%	&= \sqrt{2}  \sum_{k=1}^\infty \frac{t^k}{k!} (k-1) \Vert \QQ \Vert_F^{k-1} = \sqrt{2} \frac{e^{t\Vert \QQ\Vert_F}(t\Vert\QQ\Vert_F - 1) + 1}{\Vert \QQ \Vert_F} \, .
%\end{align*}

%\begin{figure}[!p]
%	\centering
%	\includegraphics[width=\linewidth]{frobeniusScaling.pdf}
%	\caption{Anecdotal (single simulation) error profiles with change in time $t$ and dimension. Errors calculated using Frobenius norm.}\label{fig:frob}
%\end{figure}
%
%\begin{figure}[!p]
%	\centering
%	\includegraphics[width=\linewidth]{infinityScaling.pdf}
%	\caption{Anecdotal (single simulation) error profiles with change in time $t$ and dimension. Errors calculated using infinity norm.}\label{fig:inf}
%\end{figure}

\bibliographystyle{sysbio}
\bibliography{refs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}